---
title: "Analyzing Global Inflation Rates"
output: html_notebook
---

#### This project will aim at analyzing the inflation rates globally and countrywise over time from 1980 onwards.

Loading the required packages
```{r}
require(ggplot2)
require(dplyr)
require(tidyr)

```

Reading the datset.

```{r}
library(readxl)
inflation <- read_excel("F:/PROJECTS/Datasets/IMF/inflation.xls")
View(inflation)


```
Starting with some data processing and transformations as well as cleaning.

```{r}
head(inflation)

```

The data set is quiet wide. So I will gather the dataset and make Year as a seaparate column and the inflation rates of each year separate for easy analysis and visualization of data.

```{r}
#using gather() function from tidyr

year<-c(1980:2022) #making a vector consisting of all years
year<-as.character(year)#converting to character type to use in gather()

#new dataframe which is in long format
inf<-inflation %>% gather(year,key = "Year",value="InflationRate")
inf<-na.omit(inf) #omitting NA values

names(inf)<-c("region","year","inflationRate")

inf$year<-as.integer(inf$year)

```

Now what we can do is easily filter the data for specific countries and make a separate data frame for them and analyze their inflation rates over time specifically and perform time series analysis.

---------------


### Inflation Rates in India

Generating a data frame for India and other major emerging and developed economies.

```{r}

changeType<-function(x1,x2) 
{
  
  as.numeric(x1); as.numeric(x2)
}

India<-filter(inf,region=="India")
India$inflationRate<-as.numeric(India$inflationRate)
India$year<-as.numeric(India$year)

China<-filter(inf,region=="China, People's Republic of")
China[1,3]<-0
changeType(c(China$year,China$inflationRate))

Ger<-filter(inf,region=="Germany")
changeType(Ger$inflationRate,Ger$year)


Japan<-filter(inf,region=="Japan")
US<-filter(inf,region=="United States")
EU<-filter(inf,region=="European Union")
UK<-filter(inf,region=="United Kingdom")
Fr<-filter(inf,region=="France")



ggplot(aes(x=year,y=inflationRate),data=India) +
  geom_point(size=2,color="orange") + 
  geom_line(color="orange") +
  scale_x_continuous(limits=c(1980,2017),breaks=seq(1980,2017,5))


  

```

